---
title: "Fish EDAS Data Organization and Move to R Server"
author: "Emma Jones"
date: "4/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readxl)
library(sf)
library(lubridate)
library(pins)
library(config)

# get configuration settings
conn <- config::get("connectionSettings")

# use API key to register board
board_register_rsconnect(key = conn$CONNECT_API_KEY,  #Sys.getenv("CONNECT_API_KEY"),
                          server = conn$CONNECT_SERVER)#Sys.getenv("CONNECT_SERVER"))
```

## Background

Fish EDAS has traditionally lived in a MS Access database (separate from the bug EDAS database) with one data enter-er (Royce Steiner) and two data QA officers (Jason Hill and Royce Steiner). In an effort to share data with more parties, the pertinent tables that will (eventually) be moved into CEDS will be cleaned up and moved to the R server. After the data is available on the R sever, it can be used to build a variety of tools. Once the official database is moved into CEDS, these tools will simply need to adjust their data source from the pinned data to the ODS environment and everything will work with a more "live" connection.

## Fish EDAS 

The money query Jason Hill has traditionally used is the "Total Number of fish" query in fish EDAS. This query pulls data from the "Stations," "FishSamps", "Fishes Master Taxa", and "Fishes" tables. Those tables are exported to excel for manipulation and pinning.

### Stations

This table holds a lot of the same information that other pinned data contains on the server except a completed WQS class field. This dataset will be cleaned up and the minimal data will be pinned to the server. One day, this data should be removed as other datasets should contain the information.

```{r stations}
fishStations <- read_excel('EDASdata/Stations.xlsx', sheet = 'Stations')
```

Fix some missing data before moving online. Fix wrong column names.

```{r fish station issues}
fishStationsFixed <- read_csv('EDASdata/Stations_EVJfixes.csv') %>% 
  mutate(HUC = as.character(HUC),
         Ecoregion = as.character(Ecoregion))
fishStations <- filter(fishStations, ! StationID %in% fishStationsFixed$StationID) %>% 
  bind_rows(fishStationsFixed)

fishStations <- rename(fishStations, 
                       'Class' = 'Northing',
                       'Special Standards' = 'Easting', 
                       'Ecoregion Name' = 'Subecoregion', 
                       'VAHU6' = 'DCRWSHED',
                       'HUC8' = 'HUC',
                       'Catchment Acreage' = 'Catchment  Acreage',
                       'Station Comments' = 'Comments')
```

Organize all watershed area information available from Probmon datasets and conserve va project.

```{r organize watershed area as available}
source('organizeWatershedArea.R')

```




Pin to server.

```{r pin stations}
pin(fishStations, Description = paste0('Station information from Fish EDAS, updated ', Sys.Date()), board = 'rsconnect')
```


### FishSamps

I had to manually force the Start Time and End Time fields to "Time" format in Excel after export from Access.

```{r fishSamps}
fishSamps <-  read_excel('EDASdata/fishSamps.xlsx', sheet = 'FishSamps') %>% 
  rename('FishSampID' = 'FSampID',
         'RepNum' = 'FRepNum', 
         `Collection Date` = 'CollDate', 
         'Field Team' = 'Investigators',
         'Reach Length' = 'StWidMax',
         'FishSamps Comments' = 'Comments') %>% 
  mutate(`Collected By` = NA, 
         Taxonomist = NA,
         `Start Time` = as.character(`Start Time`),
         `End Time` = as.character(`End Time`)) %>% 
  separate(`Start Time`, c(NA, "Start Time"), sep = " ") %>% 
  separate(`End Time`, c(NA, "End Time"), sep = " ")
```

Pin to server.

```{r pin fishSamps}
pin(fishSamps, Description = paste0('FishSamps information from Fish EDAS, updated', Sys.Date()), board = 'rsconnect')
```


### Fishes

Minor renaming for consistency

```{r fishes}
fishes <-  read_excel('EDASdata/Fishes.xlsx') %>% 
  rename('FishSampID' = 'FSampID',
         'RepNum' = 'FRepNum', 
         'FinalID' = 'FFinalID',
         'Fishes Comments' = 'Comments') %>% 
  dplyr::select(-Genus) %>%  # nothing in this field so drop it to avoid confusion later
  # map anomalies to new field  
  mutate(`Anomaly Code` = Anomaly,
         Anomaly = case_when(Anomaly == 'A' ~ 'Anchor Worm',
                             Anomaly == 'B' ~ 'Black Grub',
                             Anomaly == 'D' ~ 'Deformities',
                             Anomaly == 'E' ~ 'Eroded Fins',
                             Anomaly == 'F' ~ 'Fungus',
                             Anomaly == 'H' ~ 'Leeches',
                             Anomaly == 'L' ~ 'Lesions',
                             Anomaly == 'M' ~ 'Multiple DELT Anomalies',
                             Anomaly == 'N' ~ 'No Anomalies Observed',
                             Anomaly == 'P' ~ 'Popeye',
                             Anomaly == 'R' ~ 'Parasites',
                             Anomaly == 'S' ~ 'Emaciated',
                             Anomaly == 'T' ~ 'Tumor',
                             Anomaly == 'V' ~ 'Vouchered',
                             Anomaly == 'Y' ~ 'Yellow Grub',
                             Anomaly == 'Z' ~ 'Other',
                             TRUE ~ as.character(NA))) %>% 
  dplyr::select(FishIndex:FinalID, `Anomaly Code`, everything()) 
```


Pin to server.

```{r pin fishes}
pin(fishes, Description = paste0('Fishes information from Fish EDAS, updated ',Sys.Date()), board = 'rsconnect')
```


### Fishes Master Taxa

```{r master taxa}
fishesMasterTaxa <-  read_excel('EDASdata/Fishes_Master_Taxa.xlsx') %>% 
  rename('FinalID' = "FFinalID",
         "Tolerant" = "Tolerent", 
         "Intolerant" = "Intolerent",
         "Intolerantv2" = "Intolerentv2")
```

Pin to server.

```{r pin mastertaxa}
pin(fishesMasterTaxa, Description = paste0('Fish Master Taxa list from Fish EDAS, updated ',Sys.Date()), board = 'rsconnect')
```


## Bring fish data back in

```{r}
library(sf)

#### Fish Data
fishes <- pin_get("ejones/fishes", board = "rsconnect")
fishesMasterTaxa <- pin_get("ejones/fishesMasterTaxa", board = "rsconnect")
fishSamps <- pin_get("ejones/fishSamps", board = "rsconnect")
fishStations <- pin_get("ejones/fishStations", board = "rsconnect")
```



# Total fish query

```{r}
# cleaned up fish counts dropping different anomaly info
totalFish <- fishes %>% 
  group_by( FishSampID, FinalID) %>% 
  mutate(`Total Individuals` = sum(Individuals, na.rm = T)) %>% 
  dplyr::select(-c(`Anomaly Code`, Anomaly, Individuals, `Fishes Comments`, Hybrid, `Entered Date`)) %>% 
  ungroup() %>% 
  group_by(FishSampID, RepNum, FinalID) %>% 
  distinct(FinalID, .keep_all = T) %>% 
  left_join(dplyr::select(fishesMasterTaxa, FinalID, Genus, Species, `Genus-Species`), by = 'FinalID') %>% 
  left_join(fishSamps, by = c('FishSampID', 'RepNum')) %>% 
  left_join(dplyr::select(fishStations, StationID, StreamName, Lat, Long,
                          Order,  Class, `Special Standards`, `Catchment Area sqMile`, 
                          Basin, Ecoregion, `Ecoregion Name`), by = 'StationID') %>% 
  dplyr::select(StationID, StreamName, Lat, Long, Order,  Class, `Special Standards`, `Catchment Area sqMile`, 
                Basin, Ecoregion, `Ecoregion Name`, FishSampID, RepNum, `Collection Date`, 
                CollMeth, Duration, `Reach Length`, everything()) %>% 
  
  # drop missing location data for now
  filter(!is.na(Lat) | !is.na(Long)) %>% 
  
  st_as_sf(coords = c("Long", "Lat"),  # make spatial layer using these columns
           remove = F, # don't remove these lat/lon cols from df
           crs = 4326)
```


### Make huc8 layer

Don't need to rerun each time you update EDAS data

```{r}
# assessment <- st_read('data/GIS/AssessmentRegions_VA84_basins.shp') %>% 
#   mutate(HUC8 = str_sub(HUC10, 1, 8)) %>% 
#   group_by(HUC8) %>% 
#   summarize()
#   
# # now get huc8 names to join
# huc12 <- read_csv('data/GIS/huc8Names.csv') %>% # exported from C:/HardDriveBackup/GIS/GIS_BaseLayer_Datasets.gdb/Hydrology/VA_SUBWATERSHED_6TH_ORDER_STG
#   mutate(HUC_8 = as.character(paste0(0,HUC_8))) %>% 
#   distinct(HUC_8, .keep_all = T)
# 
# huc8 <- assessment %>% 
#   left_join(dplyr::select(huc12, HUC_8, HUC_8_NAME), by = c('HUC8'='HUC_8')) %>% 
#   dplyr::select(HUC8, HUC_8_NAME, everything())
# 
# #st_write(huc8, 'data/GIS/HUC8_EVJ.shp')
# rm(assessment);rm(huc12); rm(huc8)
```



## fish distribution by huc8

```{r}
huc8 <- st_read('data/GIS/HUC8_EVJ.shp')

huc8nameOptions <- huc8 %>% 
  st_drop_geometry() %>% 
  arrange(HUC8) %>% 
  mutate(HUC_8_NAME = as.character(HUC_8_NAME)) %>% 
  pivot_wider(names_from = HUC8, values_from = HUC_8_NAME) %>% 
  mutate(Taxa = NA) %>% 
  dplyr::select(Taxa, everything())


# organize total fish by taxa then station/year
#i = "telescope shiner"
stationBySpeciesUniqueList <- list()
taxaByHUC8 <- tibble()

notFoundInDatabase <- data.frame(Species=NA)

for(i in unique(totalFish$FinalID)){
  print(i)
  stationsWithSpecies <- filter(totalFish, FinalID == i)
  if(nrow(stationsWithSpecies) > 0){
    speciesCount <- suppressMessages(
      stationsWithSpecies %>%
      arrange(`Collection Date`) %>% # rearrange to make Years Collected make sense
      group_by(StationID) %>%
      summarise(`Total Count` = sum(`Total Individuals`, na.rm = T),
                `Years Collected` = paste0(year(`Collection Date`),  collapse = ", ")) %>%
      left_join(stationsWithSpecies %>% ungroup() %>%
                  dplyr::select(StationID, StreamName, Lat, Long, Order, `Catchment Area sqMile`, Class, `Special Standards`) %>%
                  distinct(StationID, .keep_all = T),
                by = 'StationID') %>%
      st_as_sf(coords = c("Long", "Lat"),  # make spatial layer using these columns
               remove = T, # don't remove these lat/lon cols from df
               crs = 4326) %>% 
      st_intersection(huc8) %>% 
      rename('Total Count' = 'Total.Count',
             'Years Collected' = 'Years.Collected', 
             'Catchment Area sqMile' = 'Catchment.Area.sqMile',
             'Special Standards' = 'Special.Standards') %>% 
      mutate(Taxa = i))
    
    huc8summary <- bind_rows(huc8nameOptions,
                             speciesCount %>%
                               ungroup() %>% st_drop_geometry() %>%
                               #distinct(HUC8, .keep_all = T) %>% 
                               group_by(Taxa, HUC8) %>% 
                               summarise(`Total Count HUC8` = as.character(sum(`Total Count`, na.rm = T))) %>%  # character to match wide format
                               arrange(HUC8) %>% 
                               pivot_wider(names_from = HUC8, values_from = `Total Count HUC8` ) )
    
      
    stationBySpeciesUniqueList[[i]] <- speciesCount
    taxaByHUC8 <- bind_rows(taxaByHUC8, huc8summary)
  }else{
    notFoundInDatabase <- rbind(notFoundInDatabase,i)
  }
}
notFoundInDatabase <- na.omit(notFoundInDatabase)
taxaByHUC8 <- bind_rows(huc8nameOptions, filter(taxaByHUC8, !is.na(Taxa)) )

write_csv(taxaByHUC8, paste0('data/taxaByHUC8_', Sys.Date(),'.csv'), na = "")

saveRDS(stationBySpeciesUniqueList, paste0('data/fishStationsUnique_',Sys.Date(), '.RDS'))
```

